---
title: "San Diego Bay permit report"
output: word_document
---

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())

# bring in some results data files from Cm_SDB_CMR and Cm_SealBeach directories. 

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(reshape2)
library(ggplot2)
library(RMark)
library(ggridges)

SE <- function(x) sqrt(var(x, na.rm = TRUE)/sum(!is.na(x)))


if (!file.exists(paste0('data/SDB_report_', Sys.Date(), '.csv'))){
  source("Sd_Bay_Report.R")
} else {
  turtle.haplo.SDB <- read.csv(file = paste0('data/SDB_report_', Sys.Date(), '.csv'))
}

reporting.date.begin <- as.Date("2019-10-01")
reporting.date.end <- as.Date("2020-09-30")

turtle.haplo.SDB %>%
  mutate(SCL.regress = ifelse(is.na(SCL) & !is.na(CCL), 1, 0),
         CCL.regress = ifelse(is.na(CCL) & !is.na(SCL), 1, 0),
         Mass.regress = ifelse(is.na(Weight) & !is.na(SCL), 1, 0),
         SCL = ifelse(SCL.regress == 1,
                      (CCL - 0.64)/1.06, SCL),
         CCL = ifelse(CCL.regress == 1,
                      0.64 + 1.06 * SCL, CCL),
         Weight = ifelse(Mass.regress == 1,
                         exp(0.96 + 0.04 * SCL),
                         Weight),
         Date = as.Date(paste(Yr, Mo, Da, sep = "-"),
                        format = '%Y-%m-%d')) -> turtle.haplo.SDB

turtle.haplo.SDB %>% filter(Date >= reporting.date.begin &
                              Date <= reporting.date.end) -> new.captures

new.captures %>% group_by(NMFS_Tag) %>%
  summarise(Sex = first(Sex),
            CCL = first(CCL),
            Weight = first(Weight),
            Date = first(Date),
            SCL = first(SCL)) -> unique.new.captures

```

# 1. What progress did you make toward meeting your objectives this year? Summarize what you did and if and how you met your objectives. List citations for any reports, publications, and presentations from this reporting period. We may request electronic copies. 

## Research Summary

Field operations were undertaken under NMFS Scientific Research Permit 18238 in San Diego Bay between ```r min(new.captures$Date)``` and ```r max(new.captures$Date)```.  During the permit year, ```r length(unique(new.captures$Date))``` field outings were made to San Diego Bay with a total of ```r length(unique(new.captures$NMFS_Tag))``` green turtle interactions.  Also, at the San Gabriel River/Seal Beach National Wildlife Refuge ...  

Biological and morphological data (lengths, mass and body depth) were collected and recorded on each of the captured animals.  Animals were checked for existing Inconel and PIT tags and new animals received both an inconel flipper tag and a PIT tag.  Skin samples for genetics and/or stable isotope analysis were collected from each of the animals and blood for hormone and potential contaminant analysis was collected from a majority of the animals. Esophageal lavage was not conducted.  

Of the ```r nrow(new.captures)``` interactions in San Diego Bay with ```r length(unique(new.captures$NMFS_Tag))``` animals, ```r nrow(filter(unique.new.captures, Sex == "F"))``` of these were classified as females, ```r nrow(filter(unique.new.captures, Sex == "M"))``` were classified by males, and ```r nrow(filter(new.captures, Sex == "U"))``` were juveniles with an unknown sex (Table 1), based upon the mean nesting size eastern Pacific green turtles.  X were new animals not previously captured (See Table 1 with *).  This group of previously untagged turtles likely consists of recent recruits into San Diego Bay, i.e. turtles that have not been living in the bay for long and have recently made the life-history shift from being juveniles in the pelagic ocean to larger sub-adults in the near-shore environment, as well as larger animals that have not been captured previously. Long-term recaptures are important indicators that San Diego Bay is a well-established foraging ground for juvenile turtles through adult life stages.  Capturing new juveniles suggests that animals continue to recruit into San Diego Bay.

```{r echo=FALSE}
new.captures %>% select(NMFS_Tag, Date, Sex, Weight, SCL) %>%
  arrange(by = Date) -> new.captures.table
knitr::kable(new.captures.table, format = "markdown", digits = 1,
      caption = paste0("Table 1. Summary of green turtle interactions in San Diego Bay from ", 
                       min(new.captures$Date), " to ",  max(new.captures$Date), "."))
```

  

### Abundance
During a portion of this reporting period (2018-10-01 to 2019-09-30), we conducted a short-term intense capture-recapture effort from 2019-05-07 to 2019-06-26 so that we can estimate abundance of green turtles in San Diego Bay assuming no deaths/immigration/emigration occurred during the period. 

Green turtles were caught with entanglement nets (50-100m length × 8m depth, mesh size = 40 cm stretched).  The entanglement nets used in this study have been used to capture the smallest of post-pelagic juvenile turtles at other green turtle foraging areas in Mexico (SCL ≥ 35 cm; Seminoff et al. 2003) as well as in SDB in the past (SCL ≥ 44.0 cm; Eguchi et al. 2010).  We therefore assumed that all size classes of green turtles living in SDB would be caught by these nets.  Exact locations of nets were dictated by tidal height, which restricted access to some areas during a low tide.  All nets were deployed from Boston whalers (approximately 5 m in length) with a 75-hp or 25-hp outboard motor.  The distance of the nets from the shore ranged from approximately 10 to 100 m, whereas water depths ranged from <1 to 6 m. Deployment time and latitude and longitude coordinates of two ends and the center of each deployed net were determined using a Global Positioning System (GPS) on a mobile phone. Deployment time and locations were used to compute the sampling effort. 

Capture effort was conducted during late afternoon and evening. Both boats remained near deployed nets in order to maintain close attentions to turtles and other animals that might be caught in the nets. Nets were checked at least every 45 minutes. When a turtle was found in a net, it was broght onboard, disentangled, and checked for existing tags (flipper and passive integrated transponder (PIT) tags).  If no tag was present, new tags were applied.  Curved carapace length (CCL) were measured from the nuchal notch to the posterior-most portion of the rear marginal scutes, whereas curved carapace width (CCW) was measured at the widest part of the carapace using a fiberglass measuring tape. Standard carapace length and width were measured using a caliper.  Turtles were kept onboard until all the nets were extracted to avoid recaptures of the same turtles during the evening 

#### Statistical analyses
Capture-recapture data were fitted to several statistical models to estimate abundance. We used maximum likelihood (ML) and Bayesian approaches. All models included capture probabilities (p), which were assumed constant (p.) or varying with capture occasions (p(t)).  For the ML approach, we considered "closed" and "Huggins' closed" models. The difference between the two groups of models is that models in the closed models explicitly include the number of unobserved turtles as a parameter (full likelihood), whereas those in Huggins' models do not (conditional likelihood). Huggins' approach, unlike the closed models, can accommodate individual covaraites.  We considered effects of capture occasions and body size (continuous or discrete categories) on capture probabilties. Models were compared within each group, i.e., closed or Huggins', using AICc. The most parsimonious model was used for abundance estimation. The appropriateness of the best model was determined through goodness of fit tests using UCare software (REF) via R2Ucare package (v. 1.0.0, Gimenez et al. 2017). 

For the Bayesian approach, effects of capture occasions, behavior, and variability among individuals (individual heterogeneity) on capture probability were considered.  For the individual heterogeneity model, we assumed the capature probability among individual was normally distributed in the logit scale, which was a common assumption for capture heterogeneity models (REF).  For all models, we used the data-augmentation approach of Royle and Dorazio (YR, or Royle YR, Kery and Schaub 2012).   We used vague priors for all parameters. We ran 5 independent chains of 50000 steps as a burn-in sample, which was followed by 10000 steps for posterior samples. These posterior samples were used to compute parameter estimates.  Numerical convergence of each analysis was determined using Gelman-Rubin statistic (Gelman et al. YR). Models were compared using deviance information criteria (DIC, Spiegelhalter YR).  The best estimates from two groups of the ML approach and the best estimate from the Bayesian approach were compared using the precisions of the estimates. We provide all estimates but use the most precise estimate (narrowest 95% confidence or credible interval) as our best estimate of abundance of green turtles in the south part of San Diego Bay during early summer of 2019.

All statistical analyses were conducted using the R Statistical Language (v.3.6.1; R core team 2019). The ML analyses were conducted using Mark (REF) through RMark (v. 2.2.6, Laake 2013), whereas the Bayesian analysis was conducted using JAGS (v. 4.3, Plummer 2018) through rjags (v. 4.8, Plummer 2018) in R. 

#### Results 
##### Field sampling
```{r results1, include = F, echo = FALSE, message=FALSE}
# this file is from Cm_SDB_CMR directory.

# BRING IN SIZE DATA FOR THE SAMPLING. SCL, CCL, ETC.
# this file was created using extract_data_TurtleDB_2020.R. See the script for 
# details - connect to the database using ODBC
dat.size <- read_csv(file = paste0("data/CMR2019_size_data_",
                                   Sys.Date(), ".csv"))

CH <- select(dat.size, -c("ID", "CCL_cm"))

#capt.hist <- data.frame(ch = pasty(CH))

n.turtles <- nrow(dat.size) #colSums(dat0[, 2:ncol(dat0)])
n.captures <- rowSums(CH)

```

At least one turtle was caught on ```r ncol(CH)``` days from 7 May 2019 to 26 June 2019.
We caught a total of ```r sum(n.captures)``` turtles (```r nrow(dat.size)``` unique turtles). The maximum number of turtles caught per day was ```r max(colSums(CH))``` (mean = ```r signif(mean(colSums(CH)), digits = 2)```, SE = ```r signif(SE(colSums(CH)), digits = 2)```).  The number of captures per turtle ranged from 1 to ```r max(n.captures)``` (mean = ```r signif(mean(n.captures), digits = 2)```, SE = ```r signif(SE(n.captures), digits = 2)```).  The CCL of these turtles ranged from ```r min(dat.size$CCL_cm)``` to ```r max(dat.size$CCL_cm)``` cm (mean = ```r signif(mean(dat.size$CCL_cm), digits = 3)```, SE = ```r signif(SE(dat.size$CCL_cm), digits = 2)```).  

##### Statistical analysis
Capture and non-capture of the ```r nrow(dat.size)``` turtles during the study period were used to estimate the total abundance within the south SDB. We used body size (CCL) as a covariate that potentially affected the capture probabilities. CCL was used as a continuous variable or binary categorical variables, where the effect of cut off value was tested at 70, 80, and 90 cm. 

```{r RMark_1, include = F, echo=F, message=F, cache=TRUE}
# load results from the most recent run here:
load('RData/Mark_Analysis_output_2019-09-24.RData')

closed.model.table <- arrange(model.list.closed$model.table, rownames(model.list.closed$model.table))

closed.Nhats <- data.frame(model = closed.model.table$model,
                           AICc = closed.model.table$AICc,
                           weight = closed.model.table$weight,
                           Nhat = unlist(lapply(model.list.closed, 
                                                function(x) x$results$derived$`N Population Size`$estimate)),
                           se = unlist(lapply(model.list.closed, 
                                                function(x) x$results$derived$`N Population Size`$se)),
                           lcl = unlist(lapply(model.list.closed, 
                                                function(x) x$results$derived$`N Population Size`$lcl)),
                           ucl = unlist(lapply(model.list.closed, 
                                                function(x) x$results$derived$`N Population Size`$ucl))) %>%
  arrange(by = AICc) %>%
  mutate(DeltaAIC = AICc - min(AICc))

Huggins.model.table <- arrange(model.list.Huggins$model.table, rownames(model.list.Huggins$model.table))
Huggins.Nhats <- data.frame(model = Huggins.model.table$model,
                           AICc = Huggins.model.table$AICc,
                           weight = Huggins.model.table$weight,
                           Nhat = unlist(lapply(model.list.Huggins, 
                                                function(x) x$results$derived$`N Population Size`$estimate)),
                           se = unlist(lapply(model.list.Huggins, 
                                                function(x) x$results$derived$`N Population Size`$se)),
                           lcl = unlist(lapply(model.list.Huggins, 
                                                function(x) x$results$derived$`N Population Size`$lcl)),
                           ucl = unlist(lapply(model.list.Huggins, 
                                                function(x) x$results$derived$`N Population Size`$ucl))) %>%
  arrange(by = AICc) %>%
  mutate(DeltaAIC = AICc - min(AICc))


```


```{r model_averaging, echo=FALSE, message=FALSE}
# model average the abundance estimates: 
# code from here: http://www.phidot.org/forum/viewtopic.php?f=21&t=952 
# Thanks Jeff Laake!

# model average Nhat value
num.models <- nrow(model.list.Huggins$model.table)
estimate <- vector("numeric",length=num.models)
se <- vector("numeric",length=num.models)
# The actual model number is the row number for the model.table
model.numbers <- as.numeric(row.names(model.list.Huggins$model.table))

for(i in 1:num.models){
  # For each model extract the derived parameter valuess and their se
  x <- model.list.Huggins[[model.numbers[i]]]$results$derived
  estimate[i] <- x$`N Population Size`$estimate
  se[i] <- x$`N Population Size`$se
}

# Call model.average using the list structure which includes estimate, weight and vcv list in this case
Nhat.Huggins.avg <- model.average(list(estimate = estimate,
                                       weight=model.list.Huggins$model.table$weight,
                                       se=se))

```

Goodness of fit tests indicated that the model (closed model) was appropriate for the data where no test indicated a significant departure from the expected numbers of recaptures (results not shown). No goodness of fit tests are available for Huggins' closed models. 

Among the closed models compared, the simplest (p(.)) was considered the most parsimonious model according to AICc values (Table 1). Estimated abundance from the model was ```r signif(Nhat.closed[[1]]$estimate, 3)``` (SE = ```r signif(Nhat.closed[[1]]$se, digits = 3)```, 95% CI = [```r signif(Nhat.closed[[1]]$lcl, digits = 3)``` - ```r signif(Nhat.closed[[1]]$ucl, digits = 3)```]. 

For the Huggins' models, the model with a covariate of size classes separated by 80 cm was the most parsimonious model.  The top 5 models, however, were within deltaAIC < ```r signif(max(model.list.Huggins$model.table$DeltaAICc[1:5]), digits = 2)```, indicating not much difference among these five models. The estimated abundance from the most parsimonious Huggins' model (p(size.cat2) was ```r signif(Huggins.Nhats[1, "Nhat"], digits = 3)``` (SE = ```r signif(Huggins.Nhats[1, "se"], digits = 3)```) with its 95% CI [```r signif(Huggins.Nhats[1, "lcl"], digits = 3)``` - ```r signif(Huggins.Nhats[1, "ucl"], digits = 3)```]. Abundance estimates from the Huggins' models were similar (Table 1).  Using the model averaging, the estimated abundance was ```r signif(Nhat.Huggins.avg$estimate, 3)``` (SE = ```r signif(Nhat.Huggins.avg$se, 3)```, approximate 95% CI = [```r signif(Nhat.Huggins.avg$estimate - 2 * Nhat.Huggins.avg$se, 3)``` - ```r signif(Nhat.Huggins.avg$estimate + 2 * Nhat.Huggins.avg$se, 3)```]).

```{r Bayes_1, include = F, echo=F, message=F, cache=TRUE}
# rather than running the whole thing, load the results from the most recent run:
Bayes.out <- readRDS(file = 'RData/Bayes_Analysis_output_2019-09-24.rds')
Bayes.out$summary$DeltaDIC <- Bayes.out$summary$DIC - min(Bayes.out$summary$DIC)
```

For the Bayesian approach, the individual random effects model (Mh) was the best model according to the deviance information criteria (DIC; Table 2). The estimated abundance (mean) was ```r signif(Bayes.out$summary[1, 'N'], digits = 3)``` (95% PI = [```r Bayes.out$summary[1, 'N_lcl']``` - ```r Bayes.out$summary[1, 'N_ucl']```]).  The point estimate from the Mh was significantly greater than the other estimates, including those from closed and Huggins' closed models.  

The closed model with the maximum likelihood approach provided the most precise estimate of abundance. (47.1, SE = 11.2, 95% CI = [33.9 - 82.2]).  During the early summer of 2017, we conducted a similar study to estimate the abundance.  The analysis of the data indicated the estimated abundance of 48 (SE = 7.03, 95% CI = [38.9 – 68.7]).  The two estimates of abundance are not statistically different. 


<!-- ```{r Bayes_2, include = F, echo=F, message=F, cache=TRUE} -->
<!-- #Mh.samples <- runjags::combine.mcmc(bayes.out.Mh$sample) -->
<!-- Mh.samples.df <- as.data.frame(runjags::combine.mcmc(Bayes.out$Mh$samples)) -->

<!-- p.N <- ggplot(data = Mh.samples.df) +  -->
<!--   stat_density(aes(x= N), geom = "line", -->
<!--                size = 1.5) +  -->
<!--   geom_segment(aes(x = Bayes.out$Mh$q2.5$N, xend = Bayes.out$Mh$q97.5$N, -->
<!--                    y = 0.0, yend = 0.0), -->
<!--                size = 1.2, arrow = arrow(ends = "both")) +  -->
<!--   geom_point(aes(x = Bayes.out$Mh$q50$N, y = 0.0), -->
<!--              size = 5) + -->
<!--   xlab("") + ylab("Density") -->
<!--   #geom_density(aes(x = N), size = 1.0) -->
<!-- ``` -->


<!-- ```{r plot_postN_Mh, echo=FALSE, fig.cap="Marginal posterior of abundance (N). The arrow indicates the 95% posterior interval and the filled circle indicates the median."} -->
<!-- p.N -->
<!-- ``` -->

### Change in body size
Because our capture-recapture effort started in 1990, body size of captured turtles in the reporting period (from 1 October 2018 to 30 September 2019) was examined in the context of the entire dataset. Missing SCL (cm) or CCL (cm) was predicted using the regression equation in Eguchi et al. (2010: CCL = 0.64 + 1.06SCL, n = 211, r2 = 0.99). Missing body mass (kg) was also predicted using the regression equation in Eguchi et al. (2010: ln(Mass) = 0.96 + 0.04SCL, n = 177, r2 = 0.96).

Standard carapace length of captured turtles during the reporting period ranged from ```r min(unique.new.captures$SCL, na.rm = T)``` cm to ```r max(unique.new.captures$SCL, na.rm = T)``` cm (mean = ```r signif(mean(unique.new.captures$SCL), 3)``` cm, SE = ```r signif(SE(unique.new.captures$SCL), 2)``` cm, n = ```r nrow(unique.new.captures)```, whereas for the entire dataset mean = ```r signif(mean(turtle.haplo.SDB$SCL, na.rm = T), 3)``` cm and SE = ```r signif(SE(turtle.haplo.SDB$SCL), 3)``` cm (n = ```r nrow(turtle.haplo.SDB)```. 

```{r echo=FALSE, warning=FALSE}
estim.SCL <- filter(turtle.haplo.SDB, SCL.regress == 1)

p1 <- ggplot()  +
  geom_path(data = turtle.haplo.SDB,
            aes(x = Date, y = SCL, group = NMFS_Tag)) + 
  geom_point(data = turtle.haplo.SDB,
             aes(x = Date, y = SCL)) +
  geom_rect(aes(xmin = reporting.date.begin, 
                xmax = reporting.date.end,
                ymin = 40, ymax = 110), 
            alpha = 0.6,
            fill = "darkgray") +
  ylab("SCL (cm)") + xlab("") + 
  theme(axis.text = element_text(size = 12))


  # geom_point(data = estim.CCL, 
  #            aes(x = Date, y = SCL),
  #            color = "red")

```

```{r Change_SCL, fig.cap= "Figure 1. Changes in SCL (cm) of green turtles captured in San Diego Bay from 1990 to present. The current reporting period is shown as the shaded region. Lines connect the recaptured individuals."}

p1
```


Distributions of SCL appeared to change over time, where a greater proportion of smaller individuals (< 80 cm) was found in the last 5 years than before (Figure 2). This may be a result of the increasing population.

```{r change_size, echo = FALSE, warning=FALSE, message=FALSE}
turtle.haplo.SDB %>% mutate(Year.Group = if_else(Yr>=2015, "2015-2019",
                                                 if_else(Yr>=2010 & Yr<=2014, "2010-2014",
                                                         if_else(Yr>=2005 & Yr <= 2009, "2005-2009",
                                                                 if_else(Yr>=2000 & Yr <= 2004, "2000-2004", "1990-1999"))))) -> Clean_data2

Clean_data2 %>% group_by(NMFS_Tag) %>% 
  summarize(firstCap = first(Date),
            firstGroup = first(Year.Group),
            firstSCL = first(SCL)) -> firstcap_only

p2 <- ggplot(data=firstcap_only, 
       aes(x = firstSCL, y = firstGroup), na.rm =T)+
  geom_density_ridges2(fill="forestgreen",
                       jittered_points = T,
                       position = position_points_jitter(height = 0),
                       alpha = 0.5)+
  labs(main="Change in size distribution", x="SCL(cm)", y="")

```

```{r changeSCL, echo = FALSE, warning = FALSE, fig.cap="Figure 2. Changes in SCL (cm) distribution of green turtles captured in San Diego Bay in 5-year segments from 1990 to present. Data points for each 5-year segment are shown as a rug plot."}

p2
```


### Genetics
Mitochondrial DNA sequencing is conducted on each individual turtle sampled from San Diego Bay.  Data analysis indicates that the animals are of Eastern Pacific origin with Islas Revillagigedos and Michoacan, Mexico contributing a majority of animals to the site (Dutton et al. 2019). 

### Stable Isotopes
Skin and blood samples for stable isotope analysis continue to be taken for long term monitoring and analysis. Currently, analyses of these samples are ongoing. Once analyses are completed, data will be compared to that of San Diego Bay, reported in Schumacher et al. in prep.

### Satellite Telemetry

As part of the on-going ecological study of East Pacific green turtles in San Diego Bay, NOAA has been collaborating with Unified Port of San Diego and U.S. Navy to determine movements and habitat use of the turtles. Past studies (funded by the Port) that were conducted prior to the closure of the Chula Vista power plant in 2010 have indicated that green turtles in the bay remain in the southern part of San Diego Bay (MacDonald et al. 2012, MacDonald et al. 2013). Because of the expected changes in the habitat and their effects on the green turtles following the closure of the power plant in 2010, studies on their movements are ongoing.

#### Results


## List titles of reports, publications, etc. resulting from this reporting period. 
Dutton PH et al. (2019)
Eguchi et al. (accepted)



# 2. Summarize how animals reacted to specific procedures. Include normal and abnormal responses of target and non-target animals. Where possible, provide quantitative data and estimate the proportion of animals (%) that had those reactions. 

During the course of the permit period, we did not encounter any unusual problems or unforeseen effects on the animals or in our research efforts. The tagging and sampling methods are standard practice and no adverse reactions were observed. All interacted animals have had good vigor upon release and many have been recaptured at a later date indicating that the animals remain in the area (e.g. San Diego Bay) after capture. 

The permitted activities appeared to have minimal effects on the animals. Most animals were calm after capture and during the sampling procedures. During capture, close attention is paid to the animals' activity to ensure that stress is minimized. A tarp was placed underneath the animal and box was used to contain the animal to prevent injuries.  A cool, wet cloth was used to wipe the carapace periodically to reduce the possibility of overheating.  During day time captures, an ez-up shade structure was used to provide shade to minimize exposure to the sun/heat.  Sampling events and transmitter applications did not produce any unforeseen responses.

# 3. Explain your efforts to conduct follow-up monitoring. Report your findings.

This is a long-term monitoring project and our field research occurs on an annual basis.  Recaptured animals are examined in order to determine any effects from previous sampling, tagging or transmitter events.  Upon recapture, all sampling sites have been healed and there is little evidence of a transmitter attachment once the device has been shed.  In addition, we photograph each of the animals to document injuries they may have sustained over time.  Inconel and PIT retention is good at our field site.  It is rare for us to encounter an animal that has lost a tag.  

# 4. Did serious injuries or mortalities occur or did you take a protected species you were not permitted to take? If so, and you already submitted an incident report, please briefly describe the event here and refer to the incident report.

We did not have any instances during the season in which animals were unintentionally injured or killed.

# 5. Describe any other problems encountered during this reporting period and steps taken or proposed to resolve them. Examples include equipment failure, weather delays, safety issues, and unanticipated effects to habitats or other species. 

No problems were encountered.

# 6. What efforts did you make to coordinate with the applicable NMFS Regional Office(s) and collaborate with other researchers? How did you collaborate (for example, avoiding field work at the same time or working together on the same animals, sharing vessels, sharing data)? 

Phone calls were made to U.S. Fish & Wildlife and California Department of Fish and Game Agents 24 hours prior to research outings.  In addition, the Port of San Diego and Navy were notified of the research outings and Navy biologists are often on hand to help during capture efforts.

#Literature Cited:
Cooch E, White G. Program MARK. Available from http://www.phidot.org  
DUTTON DL, DUTTON PH. 1999. Accelerated growth in San Diego Bay green turtles?  Proceedings of the 17th Annual Symposium on Sea Turtle Biology and Conservation   NOAA Technical Memorandum NMFS-SEFSC-415:164-165.

DUTTON PH, BALAZS GH, LEROUX RA, MURAKAWA SKK, ZARATE P, SARTI-MARTINEZ L. 2008. Genetic Composition of Hawaiian green turtle foraging aggregations: mtDNA evidence for a distinct regional population. Endangered Species Research 5:37-44.

EGUCHI T, SEMINOFF JA, LEROUX RA, DUTTON PH, DUTTON DL. 2010. Abundance and survival rates of green turtles in an urban environment: coexistence of humans and an endangered species. Marine Biology 157:1869-1877.

FELGER RS, CLIFFTON K, REGAL PJ. 1976. Winter dormancy in sea turtles: Independent discovery and exploitation in the Gulf of California by two local cultures. Science 191:283-285.

Gelman A, Carlin JB, Stern HS, Dunson DB, Vehtari A, Rubin DB. 2014. Bayesian data analysis. CRC Press
Kéry M, Schaub M. 2012. Bayesian population analysis using WinBUGS. Academic Press

Laake JL 2013. RMark: An R Interface for Analysis of Capture-Recapture Data with MARK. AFSC Processed Rep 2013-01, 25p. Alaska Fish. Sci. Cent., NOAA, Natl. Mar. Fish.  Serv., 7600 Sand Point Way NE, Seattle WA 98115.

MACDONALD BD, LEWISON RL, MADRAK SV, SEMINOFF JA, EGUCHI T. 2012. Home range of East Pacific green turtle Chelonia mydas in a highly urbanized temperate foraging ground. Marine Ecology Progress Series 461:211-221.

MACDONALD BD, MADRAK SV, LEWISON RL, SEMINOFF JA, EGUCHI T. 2013. Fine scale diel movement of the east Pacific green turtle, Chelonia mydas, in a highly urbanized foraging environment. Journal of Experimental Marine Biology and Ecology 443:56-64.

MADRAK SV, LEWISON RL, SEMINOFF JA, EGUCHI T. 2016. Characterizing response of East Pacific green turtles to changing temperatures: using acoustic telemetry in a highly urbanized environment. Animal Biotelemetry 4:22

MCDONALD D, DUTTON P, MAYER D, MERKEL K. 1994. Review of the green turtles of South San Diego Bay in relation to the operations of the SDG&E South Bay Power Plant. Doc 94-045-01. Prepared for San Diego Gas & Electric Co., C941210311. San Diego, CA. 

MOON D-Y, MACKENZIE DS, OWENS DWM. 1997. Simulated hibernation of sea turtles in the laboratory: I. Feeding, breathing frequency, blood pH, and blood gases. The Journal of Experimental Zoology 278:372-380.
Plummer M 2016. rjags: Bayesian Graphical Models using MCMC. R package version 4-6. https://CRAN.R-project.org/package=rjags

Plummer M 2017 JAGS. Available from http://mcmc-jags.sourceforge.net/ 

Royle JA, Dorazio RM. 2008. Hierarchical modeling and inference in ecology. Academic Press

SEMINOFF, J. A. 2000. The biology of the East Pacific green turtle (Chelonia mydas agassizii) at a warm temperate foraging area in the Gulf of California, Mexico. Ph. D. Dissertation. University of Arizona, Tucson. 248 pp.

SEMINOFF JA, JONES TT, RESENDIZ A, NICHOLS WJ, CHALOUPKA MY. 2003. Monitoring green turtles (Chelonia mydas) at a coastal foraging area in Baja California, Mexico: multiple indices describe population status. Journal of Marine Biology 893:1335-1362.

STINSON ML 1984 Biology of sea turtles in San Diego Bay, California, and in the northeastern Pacific Ocean. M.S. Thesis. San Diego State University. San Diego, CA. 578 pp. 


